---
title: "Central Park Squirrels | Tidy Tuesday | 2023-05-23"
format: html
editor: visual
---

```{r}
library(tidytuesdayR)
library(stringr)
library(dplyr)
library(osmdata)
library(sf)
library(ggplot2)
library(png)

library(showtext)
library(grid)
library(ggimage)
library(magick)
```

## Data import

```{r}
#| message: false
#| warning: false
squirrels_data <- tt_load("2023-05-23")$squirrel_data

colnames(squirrels_data) <- colnames(squirrels_data) %>% 
  str_to_lower() %>% 
  str_replace_all(pattern = " ", replacement = "_")
```

## Data viz

### Create a map of Central Park, NYC City

Use the `{osmdata}` to extract and use data from OpenStreetMap (OSM).

Two features are considered here:

- water: basins and lakes in Central Park
- highway: paths, footways, cycleways, etc.

```{r}
# available_tags("water") # available possibilities for the feature 'water'
# str_c(available_tags("water"), collapse = "', '")

central_park_boundary <- c(
  left = min(squirrels_data$x) - 0.001,
  bottom = min(squirrels_data$y) - 0.001,
  right = max(squirrels_data$x) + 0.001,
  top = max(squirrels_data$y) + 0.001
)

spatial_info_water <- opq(
    bbox = central_park_boundary
  ) %>%
  add_osm_feature(key = "water",
                  value = c("reservoir", "pond", "basin")) %>%
  osmdata_sf()

spatial_info_paths <- opq(
    bbox = central_park_boundary
  ) %>%
  add_osm_feature(key = "highway",
                  value = c(
                    "primary", "secondary", "tertiary",
                    "path", "footway", "cycleway")) %>%
  osmdata_sf()
```

Plot these spatial infos with `{sf}` and `{ggplot2}`:

```{r}
map_central_park_plot <- ggplot() + 
  geom_sf(
    data = spatial_info_water$osm_polygons,
    inherit.aes = FALSE,
    fill = rgb(87, 127, 146, max = 255),
    color = NA
  ) +
  geom_sf(
    data = spatial_info_paths$osm_lines,
    inherit.aes = FALSE,
    alpha = 0.1,
    size = 0.1
  ) +
  theme(
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    panel.grid = element_blank(), 
    axis.title = element_blank(), 
    axis.text = element_blank(), 
    axis.ticks = element_blank()
  )

map_central_park_plot
```

### Add the location of squirrels in Central Park, NYC

```{r}
map_squirrels_plot <- map_central_park_plot +
  geom_point(
    data = squirrels_data,
    mapping = aes(
    x = x, 
    y = y),
    size = 2,
    alpha = 0.2
  )

map_squirrels_plot
```

### Add the image of a squirrel and the statue of liberty

These images were extracted from Freepik, and combined.

```{r}
img_squirrel_statue <- readPNG(
  file.path(here::here(), "tidytuesday-2023-05-23/img/squirrel_statue_liberty.png")
  ) %>% 
  as.raster()

map_squirrels_image_plot <- map_squirrels_plot +
  annotation_raster(
    img_squirrel_statue, 
    xmin = central_park_boundary[["right"]] - 0.01, 
    xmax = central_park_boundary[["right"]] + 0.04,
    ymin = central_park_boundary[["bottom"]] - 0.01, 
    ymax = central_park_boundary[["top"]]) +
  coord_sf(
    xlim = c(central_park_boundary[["left"]],
             central_park_boundary[["right"]] + 0.04)
  )

map_squirrels_image_plot
```

### Add a title and a caption with an original font

Use the `{showtext}` package to register a font from Google font.

```{r}
font_add_google(
  name = "Syne Mono", 
  family = "Syne Mono"
)
showtext_auto()
```

Add it to the plot:

```{r}
map_squirrels_image_plot_title <- map_squirrels_image_plot +
  labs(
    title = "Squirrels are everywhere in\nCentral Park, NYC!",
    caption = "Data: 2018 Central Park Squirrel Census | Graphic: Margot Brd\nData (c) OpenStreetMap contributors, ODbL 1.0. https://www.openstreetmap.org/copyright"
  ) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      family = "Syne Mono",
      size = 30,
      margin = margin(
        t = 0,
        r = 0,
        b = -30,
        l = 0
      )
    ),
    plot.caption = element_text(
      hjust = 1,
      family = "Syne Mono", 
      size = 10
      ))

map_squirrels_image_plot_title
```

### Add a vintage paper in the background

```{r}
img_vintage <- readPNG(
  file.path(here::here(), "tidytuesday-2023-05-23/img/vintage_paper.png"),
  native = TRUE
)

map_squirrels_image_plot_title

grid.draw(gList(
  rasterGrob(
    img_vintage,
    width = unit(1, "npc"),
    height = unit(1, "npc")
  ),
  ggplotGrob(map_squirrels_image_plot_title)
))
```
